An HPSG account for deliminative verbal reduplication in Mandarin Chinese

### 2024-08-19 replies-reviews.tex

Finally, as we discussed in Sec. 2.3.1 in the previous version of the paper, the semantics of the reduplication has the properties of tansitoriness, holisticity and dynamicity.
This means that the use of reduplication affects the situation aspect at the sentential level.
One might argue that examples such as (\ref{ex:xiang}) and (\ref{ex:zaichang}) above convey dynamic meaning, even though stative verbs are used.
%i.e. they may be interpreted as `become similar to' and `become present' respectively.
However, we argue that the dynamic interpretation does not come from the verb but from reduplication.
These verbs are stative in a neutral context and thus, we consider the intrinsic feature of these verbs to be stative.
The dynamic interpretation only arises when they are used in specific contexts, in this case, when they are reduplicated.
The point of our examples is to show that reduplication can be applied to verbs of all verb classes.

As noted briefly in Sec. 2.2 in the previous version of the paper,
the reduplication is incompatible with an expression that quantifies the duration or the extent of the event,
 because the reduplication already contains a (abstract) quantity meaning, namely a short duration or a small extent (transitoriness),
 which cannot be measured on a concrete scale (\citealt[155]{XiaoMcEnery2004}, \citealt[333]{SuiHu2016}).
% The same can be argued for the incompatibility of reduplication and a quantized object () , as both serve to delimit the event expressed by the base verb \citep[322]{SuiHu2016}.
 %() *我看看三本书。
 This results from the properties of reduplication and does not change the verb class of the predicate.

Due to its unique features compared to the perfective aspect marker \obj{le} and the experiential aspect marker \obj{guo4},
the reduplication is the only perfective viewpoint aspect that can freely occur in future situations \citep[156--157]{XiaoMcEnery2004}.
\textit{Le} expresses a situation being realized,
while \textit{guo4} conveys a situation already experienced,
and thus, both rarely occur in future situations.
The use of reduplication does not have these constraints.
This explains the general tendency for the reduplication to appear in conditional clauses and imperative or causative contexts,
which are frequently used to describe future situations.

### 2024-08-19 chapters/ch02-Phenomenon.tex
Update 2024-08-19-15:50:11

In this section, we will first discuss the core meaning of the reduplication as well as the meaning of its different forms (\sectref{sec:core-sem}).
We will then investigate the interaction of the reduplication with verb classes (\sectref{sec:Aktionsarten}), adjuncts (\sectref{sec:adjuncts}) and aspect markers (\sectref{sec:aspM}).

@@

%%%%%%%%%
\subsubsection{Interaction with adjuncts}\label{sec:adjuncts}

@@

As mentioned in Section \ref{sec:forms}, the reduplication can only be marked by the perfective aspect marker \textit{le} but not other aspect markers.\footnote{
There is no consensus on which elements exactly are considered aspect markers in Mandarin Chinese. We only discuss the most commonly recognized ones here.
}
We believe this incompatibility to be for semantic reasons.

\citet[Ch. 4]{XiaoMcEnery2004} considered \textit{le}, \textit{guo} and reduplication to indicate perfective aspects, as they all view the situation as an inseparable whole.
The perfective aspect marker \textit{le} is compatible with reduplication while the experiential aspect marker \textit{guo} is not.
\citet[128--131]{XiaoMcEnery2004} stated that \textit{le} has the semantic feature of dynamicity, since it ``can focus on both heterogeneous internal structures and changing points'' \citep[129]{XiaoMcEnery2004}.
It can be combined with a situation with a dynamic internal structure, such as crying in (\ref{ex:cry}).
It can also co\hyp{}occur with a situation with a change at a certain time point, such as getting to know in (\ref{ex:know}).

\ea
  \ea\label{ex:cry}
    \gll Weici, Deng Lijun shangxin de ku-le san tian.\\
    for.this Deng Lijun sadly \textsc{de} cry-\textsc{pfv} three day\\ \jambox{\citep[129]{XiaoMcEnery2004}}
    \glt `For this reason, Deng Lijun cried sadly for three days.'
  \ex\label{ex:know}
    \gll Ta zhidao-le zhe shi de nei-qing.\\
    he know-\textsc{pfv} this matter \textsc{de} inside-information\\ \jambox{\citep[130]{XiaoMcEnery2004}}
    \glt `He got to know the inside information on this matter.'
  \z
\z
\textit{Le} is compatible with the reduplication, because its dynamicity can relate to not only the termination or instantiation of an event (a point of change), but also the process of the situation, just like that of the reduplication.

In comparison, the experiential aspect marker \textit{guo} cannot co\hyp{}occur with a reduplicated verb, 
because its dynamicity attributes to an ``experiential change'' \citep[148]{XiaoMcEnery2004}, 
namely that a situation has been experienced historically and that ``the final state of the situation no longer obtains'' at the reference time \citep[144]{XiaoMcEnery2004}. 
Compare (\ref{ex:guo-dyn}) and (\ref{ex:le-dyn}), \textit{guo} in (\ref{ex:guo-dyn}) suggests a change out of the state of being a soldier, 
whereas \textit{le} in (\ref{ex:le-dyn}) conveys a change into the state of being a soldier \citep[149]{XiaoMcEnery2004}.

\ea
  \ea\label{ex:guo-dyn}
    \gll ta dang-guo bing.\\
    he serve.as-\textsc{exp} soldier\\ \jambox{\citep[149]{XiaoMcEnery2004}}
    \glt `He once served as a soldier.'
  \ex\label{ex:le-dyn}
    \gll ta dang-le bing.\\
    he serve.as-\textsc{pfv} soldier\\ \jambox{\citep[149]{XiaoMcEnery2004}}
    \glt `He became a soldier.'
  \z
\z
It is clear that \textit{guo} only indicates a change at the termination of a situation and cannot express the dynamicity within a situation.
Hence, it is incompatible with the semantics of the reduplication.

Due to the holistic semantics of the reduplication, it is incompatible with imperfective aspect markers: the durative aspect marker \textit{zhe} and the progressive aspect marker \textit{zai}, as both only focus on  a part of the situation and do not view the situation as a whole \citep[Ch. 5]{XiaoMcEnery2004}.

From the illustration above, it seems that due to its semantics, reduplication can only be marked by \textit{le} but not the other aspect markers.

>>>>

In this section, we will first discuss the core meaning of the reduplication as well as the meaning of its different forms (\sectref{sec:core-sem}).
We will then investigate the interaction of the reduplication with verb classes (\sectref{sec:Aktionsarten}), aspect markers (\sectref{sec:aspM}) and other sentential components (\sectref{sec:adjuncts}) .

@@

As mentioned in \sectref{sec:forms}, the reduplication can only be marked by the perfective aspect marker \textit{le} but not other aspect markers.\footnote{
There is no consensus on which elements exactly are considered aspect markers in Mandarin Chinese. 
We only discuss the most commonly recognized ones here.
}
We believe this incompatibility to be for semantic reasons.

\citet[Ch. 4]{XiaoMcEnery2004} consider \obj{le}, \obj{guo4} and reduplication to indicate perfective aspects, as they all view the situation as an inseparable whole.
The perfective aspect marker \obj{le} is compatible with reduplication while the experiential aspect marker \obj{guo4} is not.
\citet[128--131]{XiaoMcEnery2004} stated that \obj{le} has the semantic feature of dynamicity, since it ``can focus on both heterogeneous internal structures and changing points'' \citep[129]{XiaoMcEnery2004}.
It can be combined with a situation with a dynamic internal structure, such as crying in (\ref{ex:cry}).
It can also co\hyp{}occur with a situation with a change at a certain time point, such as getting to know in (\ref{ex:know}).

\ea
  \ea\label{ex:cry}
    \gll \objex{wei4ci3}, \objex{Deng4} \objex{Li4jun1} \objex{shang1xin1} \objex{de} \objex{ku1-le} \objex{san1} \objex{tian1}.\\
    for.this Deng Lijun sadly \textsc{de} cry-\textsc{pfv} three day\\ \jambox{\citep[129]{XiaoMcEnery2004}}
    \glt `For this reason, Deng Lijun cried sadly for three days.'
  \ex\label{ex:know}
    \gll \objex{ta1} \objex{zhi1dao4-le} \objex{zhe4} \objex{shi4} de \objex{nei4-qing2}.\\
    he know-\textsc{pfv} this matter \textsc{de} inside-information\\ \jambox{\citep[130]{XiaoMcEnery2004}}
    \glt `He got to know the inside information on this matter.'
  \z
\z
\obj{Le} is compatible with the reduplication, because its dynamicity can relate to not only the termination or instantiation of an event (a point of change), but also the process of the situation, just like that of the reduplication (see \sectref{sec:core-sem}).

In comparison, the experiential aspect marker \obj{guo4} cannot co\hyp{}occur with a reduplicated verb, 
because its dynamicity attributes to an ``experiential change'' \citep[148]{XiaoMcEnery2004}, 
namely that a situation has been experienced historically and that ``the final state of the situation no longer obtains'' at the reference time \citep[144]{XiaoMcEnery2004}. 
Compare (\ref{ex:guo-dyn}) and (\ref{ex:le-dyn}), \obj{guo4} in (\ref{ex:guo-dyn}) suggests a change out of the state of being a soldier, 
whereas \obj{le} in (\ref{ex:le-dyn}) conveys a change into the state of being a soldier \citep[149]{XiaoMcEnery2004}.

\ea
  \ea\label{ex:guo-dyn}
    \gll \objex{ta1} \objex{dang1-guo4} \objex{bing1}.\\
    he serve.as-\textsc{exp} soldier\\ \jambox{\citep[149]{XiaoMcEnery2004}}
    \glt `He once served as a soldier.'
  \ex\label{ex:le-dyn}
    \gll \objex{ta1} \objex{dang1-le} \objex{bing1}.\\
    he serve.as-\textsc{pfv} soldier\\ \jambox{\citep[149]{XiaoMcEnery2004}}
    \glt `He became a soldier.'
  \z
\z
It is clear that \obj{guo4} only indicates a change at the termination of a situation and cannot express the dynamicity within a situation.
Hence, it is incompatible with the semantics of the reduplication.

Due to the holistic semantics of the reduplication, it is incompatible with imperfective aspect markers: the durative aspect marker \obj{zhe} and the progressive aspect marker \obj{zai4}, as both only focus on  a part of the situation and do not view the situation as a whole \citep[Ch. 5]{XiaoMcEnery2004}.

From the illustration above, it seems that due to its semantics, reduplication can only be marked by \obj{le} but not the other aspect markers.

%%%%%%%%%
\subsubsection{Interaction with other sentential components}\label{sec:adjuncts}

@@

A reviewer also notes that the reduplication appears frequently in imperative (example \ref{ex:redup-achi-stat1}) and conditional sentences (\ref{ex:cond}) as well as causative sentences with \obj{rang4}\slash\obj{jiao4}\slash\obj{shi3} `let/let/make' (examples \ref{ex:xiangxin}, \ref{ex:xiang}, \ref{ex:redup-achi-stat5}).
\ea\label{ex:cond}
\gll \objex{ni3} \objex{yao4shi4} \objex{gei3} \objex{wo3} \obj{zhi3-zhi} \objex{Li3-jia1} \objex{de} \objex{men2}, \objex{wo3} \objex{yuan4} \objex{gei3} \objex{ni3} \objex{gui4xia4}\\
you if give I point-point Li-family \textsc{de} door I be.willing.to give you kneel.down\\\jambox{\citep[157]{XiaoMcEnery2004}}
\glt `If you point at the door of Li's for me, I am willing to kneel down for you.'
\z
This can  be attributed to the holistic semantics of reduplication (see \sectref{sec:core-sem}).
%Due to its unique features compared to the perfective aspect marker \obj{le} and the experiential aspect marker \obj{guo4},
As illustrated in \sectref{sec:aspM}, \obj{le} expresses a situation being realized,
while \obj{guo4} conveys a situation already experienced,
and thus, both rarely occur in future situations.
The use of reduplication does not have these constraints.
This makes the reduplication the only perfective viewpoint aspect that can freely occur in future situations \citep[156--157]{XiaoMcEnery2004}.
This explains the general tendency for the reduplication to appear in conditional clauses and imperative or causative contexts,
which are frequently used to describe future situations.
